---
title: "Tidy Data, long to wide and back!"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r setup, echo=FALSE}
library(tidyverse)
library(gapminder)

```

## Your Turn 1

Use `pivot_longer()` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}
table4a

table4a %>% 
  pivot_longer(cols=2:3, 
               names_to="year", 
               values_to = "cases")
```

## Your Turn 2

1. Use `pivot_wider()` to reorganize `table2` into four columns: **country**, **year**, **cases**, and **population**. 
2. Create a new variable called `prevalence` that divides `cases` by `population` multiplied by 100000.
3. Pass the data frame to a ggplot. Make a scatter plot with `year` on the x axis and `prevalence` on the y axis. Set the colour aesthetic (`aes()`) to `country`. Use `size = 2` for the points. Add a line geom. 

```{r}
table2

table2 %>% 
  pivot_wider(names_from = type, values_from = count) %>% 
  mutate(prevalence = 100000*cases/population) 


  ggplot(aes(x=year, y = prevalence, colour = country)) + 
     geom_point(size = 2) +
      geom_line() + 
     scale_x_continuous(breaks = c(1999, 2000))
```

## Your Turn 3

Pivot longer the 5th through 60th columns of `who` into a key column: value column pair named **codes** and **n**. Then select just the `country`, `year`, `codes` and `n` variables.

```{r}
who
# dimensions of who = 7240 rows * 60 columns

who_long <- who %>% pivot_longer(cols=5:60, names_to = "codes", values_to="n")
# dimensions of who_long = 405,440 rows (7240*56 * 6 columns

```

## Your Turn 4

Use the `cases` data below. Separate the `sex_age` column into **sex** and **age** columns.

```{r}
cases <- tribble(
   ~id,     ~sex_age,
   "1",    "male_56",
   "2",  "female_77",    
   "3",  "female_49"
)

cases %>% 
separate("sex_age", into = c("sex", "age"))
```

## Your Turn 5: Challenge!

Black women and white women get breast cancer at about the same rate, but black women die from breast cancer at a higher rate than white women.

There are two CSV files in this folder containing SEER data in breast cancer incidence in white and black women.
https://ieee-dataport.org/open-access/seer-breast-cancer-data#files

For both sets of data:

1. Import the data
2. Pivot longer the columns into 2 new columns called `year` and `incidence`
3. Add a new variable called `race`. Remember that each data set corresponds to a single race.
4. Bind the data sets together using `bind_rows()` from the dplyr package. Either save it as a new object or pipe the result directly into the ggplot2 code.
5. Plot the data using the code below. Fill in the blanks to have `year` on the x-axis, `incidence` on the y-axis, and `race` as the color aesthetic. 

```{r}


______ %>% 
  ggplot(aes(x = ______, y = ______, colour = ______, group = race)) + 
    geom_point() +
    geom_line() + 
    scale_x_discrete(breaks = seq(1975, 2011, by = 5)) # show years by fives
```

***

# Take Aways

Data comes in many formats but R prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column
2. Every observation is in its own row
3. Every value is in its own cell (which follows from the above)

What is a variable and an observation may depend on your immediate goal.

