---
title: "Babynames: dplyr practice worksheet"
author: "Your name goes here"
date: "date goes here"
output: 
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=TRUE, message=FALSE, warning=FALSE}
library(babynames)
library(mosaic)
library(tidyverse)

glimpse(babynames)

```

Each year, the US Social Security Administration publishes a list of the most popular names given to babies.  In 2018, [the list](http://www.ssa.gov/oact/babynames/#ht=2) shows Emma and Olivia leading for girls, Liam and Noah for boys.

The `babynames` data table in the `babynames` package comes from the Social Security Administration's listing of the names givens to babies in each year, and the number of babies of each sex given that name. (Only names with 5 or more babies are published by the SSA.)


## 1. How many babies are represented?

```{r eval=TRUE}
babynames |>
  summarise(total = sum(n)) # a reduction verb
```

## 2. How many babies are there in each year?

```{r eval=TRUE}
babynames |> 
  group_by(year) |> 
  summarise(total= sum(n))


```
Can you plot the total number of names (and hence births) by year?
```{r}


```

## 3. How many distinct names in each year?

```{r eval=TRUE}
babynames |>
  group_by(year) |>
  summarise(name_count = n_distinct(name))
```

## 4. How many distinct names of each sex in each year?

```{r eval=TRUE}
babynames |>
  group_by(year, sex) |>
  summarise(name_count = n_distinct(name)) 


# plot it
babynames |>
  group_by(year, sex) |>
  summarise(name_count = n_distinct(name)) |> 
  ggplot()+
  aes(x= year, y= name_count, colour=sex)+
  geom_line()


```

Do boys or girls have similar number of distinct names? Plot `name_count` by `year`, coloured by `sex`

# Popularity of Jane and Mary

## 5. Track the yearly number of Janes and Marys over the years.

```{r eval=TRUE}
Result <-
  babynames |>
  filter(name %in% c("Jane", "Mary")) |> # just the Janes and Marys
  group_by(year, name) |> # for each year for each name
  summarise(count = sum(n))
```

### 6. Plot out the result

Put `year` on the x-axis and the count of each name on the y-axis.  
* *Map* the name (Mary or Jane) to the aesthetic of color.  Remember that mapping to aesthetics is always done inside the `aes()` function.
* Instead of using dots as the glyph, use a line that connects consecutive values: `geom_line()`.
* Change the y-axis label to "Yearly Births": `+ ylab("Yearly Births")`
* *Set* the line thickness to `size=2`. Remember that "setting" refers to adjusting the value of an aesthetic to a constant.  Thus, it's *outside* the `aes()` function.
* Add a vertical line to mark a year in which something happened that might relate to the increase or decrease the popularity of the name.  Example: The movie [*Whatever Happened to Baby Jane*](http://en.wikipedia.org/wiki/What_Ever_Happened_to_Baby_Jane%3F_%281962_film%29) came out in 1962.  The glyph is a vertical line: `geom_vline(xintercept=1962)`.

```{r eval=TRUE}

ggplot(Result, aes(x= year, y=count,colour = name)) +
  geom_line(size = 1.5) +
  labs(
    title = "Names over time", 
    x="Year", 
    y= "Yearly Births") +
  theme_bw()+
  geom_vline(xintercept=1962)
```



## 7. Look at the *proportion* of births rather than the count

```{r eval=TRUE}

Result2 <-
  babynames |>
  group_by(year) |>
  mutate(total = sum(n)) |>
  filter(name %in% c("Mary", "Jane")) |>
  mutate(proportion = n / total)

head(Result2)
```

Why is sex a variable in Result2? Eliminate it, keeping just the girls.


```{r}
Result2 <-
  babynames |>
  filter(sex == "F") |>
  group_by(year) |>
  mutate(total = sum(n)) |>
  filter(name %in% c("Mary", "Jane")) |>
  mutate(proportion = n / total)

```


Just as you did with count vs year, graph proportion vs year.
```{r eval=FALSE}
ggplot(Result2, aes(x= year, y=proportion,colour = name)) +
   geom_line(size = 2) +
  labs(title = "Names over time", x="Year", y= "Yearly Births")+
  geom_vline(xintercept=1962)

```

  
## 8. How has the sex ratio of babies (Females vs Males) changed over time 
```{r}



babynames |> 
  group_by(year, sex)  |> 
  summarise(total = sum(n)) |> 
  mutate(prop =total/sum(total)) |> 
  ggplot()+
  aes(x = year, y = prop, colour = sex)+
  geom_point()


```

Plot out the most popular names
```{r}
babynames %>%
  group_by(name, sex) %>%
  summarise(total = sum(n)) %>%
  arrange(desc(total))


babynames %>%
  filter(year>2016) |> 
  group_by(name, sex) %>%
  summarise(total = sum(n)) %>%
  arrange(desc(total)) %>%
  ungroup() %>%
  slice(1:20) %>%
  ggplot() +
  aes(y = fct_reorder(name, total), x = total, fill = sex) +
  geom_col() +
  theme_bw() +
  scale_x_continuous(label = scales::number)+
  labs(x=NULL, y = NULL)


```

